Irreducible Components and Dimension of the Springer Fiber of a Hook-Type Slodowy Slice

### 2024-09-01 paper.tex
up to section 4.2

\begin{lemma}\label{small_centralizer}
    \(\mathfrak{z}_{\gl_n}(f)\) is the set of upper-triangular \(X \in \gl_n\) such that for all \(i,j\),
    \[X_{i + 1,j + 1} = \frac{j(n - j)}{i(n - i)} X_{ij}.\]
\end{lemma}
\begin{proof}
    Let \(X \in \mathfrak{z}_{\gl_n}(f)\).
    Looking at the definition of \(f\) from the previous section (and zero-padding the matrices), we see that \((fX)_{ij} = i (n - i) X_{i + 1,j}\), and \((Xf)_{ij} = (j - 1)(n - (j - 1)) X_{i, j - 1}\).
    So,
    \[\forall i, j \in \{1, ..., n\}. \; i(n - i)X_{i + 1,j} = (j - 1)(n - (j - 1))X_{i, j - 1}.\]
    Taking \(j = 1\), the condition above tells us that \(\forall i \geq 2. \; X_{i, 1} = 0\).
    Taking \(j > 1\) and \(i < n\), we obtain that
    \[\forall i, j \in \{1, ..., n - 1\}. \; X_{i + 1, j + 1} = \frac{j(n - j)}{i(n - i)} X_{ij}.\]
    So, every \(X \in \mathfrak{z}_{\gl_n}(f)\) is upper triangular and satisfies the condition above.
    Conversely, it is clear that for such \(X\) we have \(fX = Xf\).
\end{proof}

\begin{lemma}\label{big_centralizer}
    \[\mathfrak{z}_{\gl_{m + n}}(F) =  \left\{\begin{pNiceArray}{ccc|ccc}
        & & & & & \vert \\
        & X & & & & b    \\
        & & & & & \vert \\
       \hline
       \text{---} & a & \text{---} &  \\
        &  & &  & Y  \\
        & & & & & 
       \end{pNiceArray} : X \in \gl_m; Y \in \mathfrak{z}_{\gl_n}(f); a,b \in \mathbb{C}^m \right\}.\]
\end{lemma}
\begin{proof}
    Let \(Z = \begin{pmatrix}Z_{11} & Z_{12}\\Z_{21} & Z_{22}\end{pmatrix} \in \mathfrak{z}_{m + n}(F)\).
    We have 
    \[\begin{pmatrix}0 & Z_{12}f\\0 & Z_{22} f\end{pmatrix} =  ZF = FZ = \begin{pmatrix}0 & 0\\fZ_{21} & fZ_{22}\end{pmatrix}.\]
    There is no restriction on \(Z_{11}\).
    The condition \(Z_{12}f = 0\) means that all but the last column of \(Z_{12}\) must be zero, and the condition \(0 = fZ_{21}\) means that all but the first row of \(Z_{21}\) must be zero.
    And the condition \(Z_{22}f = fZ_{22}\) means that \(Z_{22} \in \mathfrak{z}_{\gl_{m + n}}(f)\).
\end{proof}

These lemmas provide an explicit characterization of the Slodowy slice \(S = E + \mathfrak{z}_{\gl_{m + n}}(F)\).
\begin{corollary}\label{slodowy_slice}
    \[S = \left\{\begin{pNiceArray}{ccc|ccc}
        & & & & & \vert \\
        & X & & & & b    \\
        & & & & & \vert \\
       \hline
       \text{---} & a & \text{---} &  \\
        & & &  & e + Y\\
        &  & &  &   
       \end{pNiceArray} : a, b \in \mathbb{C}^m, X \in \gl_m, Y \in \mathfrak{z}_{\gl_n}(f)\right\}.\]
\end{corollary}

>>>>

\begin{lemma}\label{small_centralizer}
    \(\mathfrak{z}_{\gl_n}(f)\) is the set of upper-triangular \(X \in \gl_n\) such that for all \(i,j \in \{1,...,n-1\}\),
    \[X_{i + 1,j + 1} = \frac{j(n - j)}{i(n - i)} X_{ij}.\]
\end{lemma}
\begin{proof}
    Let \(X \in \mathfrak{z}_{\gl_n}(f)\).
    Looking at the definition of \(f\) from the previous section (and zero-padding the matrices), we see that \((fX)_{ij} = i (n - i) X_{i + 1,j}\), and \((Xf)_{ij} = (j - 1)(n - (j - 1)) X_{i, j - 1}\).
    So, for any \(i, j \in \{1,...,n\}\),
    \[i(n - i)X_{i + 1,j} = (j - 1)(n - (j - 1))X_{i, j - 1}.\]
    Taking \(j = 1\), the condition above tells us that for all \(i \geq 2\) we have \(X_{i, 1} = 0\).
    Taking \(j > 1\) and \(i < n\), we get that for all \(i,j \in \{1,...,n-1\}\),
    \[X_{i + 1, j + 1} = \frac{j(n - j)}{i(n - i)} X_{ij}.\]
    So, every \(X \in \mathfrak{z}_{\gl_n}(f)\) is upper triangular and satisfies the condition above.
    Conversely, it is clear that for such \(X\) we have \(fX = Xf\).
\end{proof}

\begin{lemma}\label{big_centralizer}
    \[\mathfrak{z}_{\gl_{m + n}}(F) =  \left\{\begin{pNiceArray}{ccc|ccc}
        & & & & & \vert \\
        & X & & & & b    \\
        & & & & & \vert \\
       \hline
       \text{---} & a & \text{---} &  \\
        &  & &  & Y  \\
        & & & & & 
       \end{pNiceArray} : X \in \gl_m; Y \in \mathfrak{z}_{\gl_n}(f); a,b \in \mathbb{C}^m \right\}.\]
\end{lemma}
\begin{proof}
    Let \(Z = \begin{pmatrix}Z_{11} & Z_{12}\\Z_{21} & Z_{22}\end{pmatrix} \in \mathfrak{z}_{m + n}(F)\).
    We have 
    \[\begin{pmatrix}0 & Z_{12}f\\0 & Z_{22} f\end{pmatrix} =  ZF = FZ = \begin{pmatrix}0 & 0\\fZ_{21} & fZ_{22}\end{pmatrix}.\]
    There is no restriction on \(Z_{11}\).
    The condition \(Z_{12}f = 0\) means that all but the last column of \(Z_{12}\) must be zero, and the condition \(0 = fZ_{21}\) means that all but the first row of \(Z_{21}\) must be zero.
    Finally, the condition \(Z_{22}f = fZ_{22}\) means that \(Z_{22} \in \mathfrak{z}_{\gl_n}(f)\).
\end{proof}

Taken together, the previous lemmas provide a nice characterization of the Slodowy slice \(S = E + \mathfrak{z}_{\gl_{m + n}}(F)\).
\begin{corollary}\label{slodowy_slice}
    \[S = \left\{\begin{pNiceArray}{ccc|ccc}
        & & & & & \vert \\
        & X & & & & b    \\
        & & & & & \vert \\
       \hline
       \text{---} & a & \text{---} &  \\
        & & &  & e + Y\\
        &  & &  &   
       \end{pNiceArray} : a, b \in \mathbb{C}^m, X \in \gl_m, Y \in \mathfrak{z}_{\gl_n}(f)\right\}.\]
\end{corollary}
